\vspace{-2.75cm}Idea Brief\vspace{-0.5cm}

### 2022-11-13 Extra.tex
idea brief

>>>>

----------
Supply side focus thus far (how donor behavior has been affected).
Lots of academic attention on supply side provision of aid given China's entrance (e.g., \cite{dreher2018}, \cite{dreher2015}), but few explorations of the demand side after China has become a prevalent donor. Also research on how Chinese aid affects recipients in various ways (e.g., \cite{martorano2020}, \cite{bader2015}).
This paper seeks to address this shortage in the literature, aiming to understand how recipient countries choose between the vast number of credible donors. Relevant literature: \cite{kilama2016a}, \cite{hernandez2017}, \cite{li2017a}, \cite{isaksson2018}, \cite{humphrey2019}, \cite{broich2017a}.

### 2022-11-13 Paper.tex
idea brief

\documentclass{article}
\usepackage[utf8]{inputenc}
\title{\vspace{-2.75cm}634\vspace{-0.5cm}}
\author{Nicholas Ray}
\date{\vspace{-0.30cm}October 31 2022\vspace{-1cm}}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools,amssymb,amsthm}
\usepackage{setspace}
\usepackage{tikz}
\usepackage{sgame}
\doublespacing
\usepackage[backend=biber, style=authoryear, maxbibnames=99, uniquelist=false]{biblatex}
\renewbibmacro{in:}{}
\renewbibmacro*{volume+number+eid}{%
  \printfield{volume}
  \setunit*{\addnbthinspace}
  \printfield{number}
  \setunit{\addcomma\space}
  \printfield{eid}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\AtEveryBibitem{
  \clearfield{issn}
  \clearfield{month}
  \clearfield{urlyear}
  \clearlist{language}
  \clearfield{note}
  \ifentrytype{online}{}{
    \clearfield{url}
  }
}
\addbibresource{ForeignAidBib.bib}
\begin{document}
\maketitle
\section*{Introduction}
%Look at Dreher's book, Kidd's international formal theory book, Hernadez 2017 paper, etc.
BRI (\cite{dreher2022}), Globalization (papers from week 6)?

Lots of academic attention on supply side provision of aid given China's entrance (e.g., \cite{dreher2018}, \cite{dreher2015}), but few explorations of the demand side after China has become a prevalent donor. Also research on how Chinese aid affects recipients in various ways (e.g., \cite{martorano2020}, \cite{bader2015}).
This paper seeks to address this shortage in the literature, aiming to understand how recipient countries choose between the vast number of credible donors. Relevant literature: \cite{kilama2016a}, \cite{hernandez2017}, \cite{li2017a}, \cite{isaksson2018}, \cite{humphrey2019}, \cite{broich2017a}.

Research Question. Given the literature, the decision-making of recipient countries no longer seems so trivial (i.e., why are countries choosing to receive aid money from China?). The literature implies that there should be a healthy mix of aid donor money given the effects shown, implying that there should be some strategy present in recipient countries.

\section*{Theory}
In an attempt to model the effects of having more than one donor on recipient behavior, especially when donors display heterogeneity in their loaning practices, I formalize a bargaining interaction between a recipient country and two types of donors. Figure 1 illustrates the extensive-form of the interaction and can be found in the appendix.

@@

The first donor's offer is largest, though, when the severity of recipient need is low (high $\delta$) and the second donor is anticipated to implement clear conditions (high $\gamma$). This implies the comparative statics that as severity of need increases (i.e., $\delta$ gets smaller) or clarity of conditionality decreases (i.e., $\gamma$ gets smaller), loan value decreases for the recipient. Theoretically, loan value to the recipient can be manipulated by either altering the amount of money loaned or by changing the conditions to be imposed.

The above analysis produces the following hypotheses, conditional on the presence of a diverse set of donors:

\begin{align*}
    H_{1_a}&:\;\text{An increase (decrease) in severity of recipient need will decrease (increase) the amount of money}\\
    &\;\text{offered via loans}\\
    H_{1_b}&:\;\text{An increase (decrease) in severity of recipient need will increase (decrease) the stringency of loan}\\
    &\;\text{conditions}\\
    H_{2_a}&:\;\text{An increase (decrease) in the clarity of loan conditionality will increase (decrease) the amount of}\\
    &\;\text{money offered via loans}\\
    H_{2_b}&:\;\text{An increase (decrease) in the clarity of loan conditionality will decrease (increase) the stringency}\\
    &\;\text{of loan conditions}\\
\end{align*}

Implications: the more unofficial, ambiguous loans that a country receives from China, the less competitive offer they will receive from more official, Western flows. The more ODA-like, official loans they receive from China, the better offers they will receive from Western institutions.

\section*{Operationalization}
Western loans (bilateral and multilateral?) as first donor, China as second with the variation in officialness in aid conditionality.

Data \& Variables
Aiddata for Chinese aid, OECD data for Western aid (World Bank or IMF?), GDP per capita part of need but need a different measure to get at temporal shocks/crises.

Research Design

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\title{\vspace{-2.75cm}Idea Brief\vspace{-0.5cm}}
\author{Nicholas Ray}
\date{\vspace{-0.30cm}October 31 2022\vspace{-1cm}}
\usepackage[margin=1in]{geometry}
\usepackage{mathtools,amssymb,amsthm}
\usepackage{setspace}
\usepackage{tikz}
\usepackage{sgame}
\doublespacing
\usepackage[backend=biber, style=authoryear, maxbibnames=99, uniquelist=false]{biblatex}
\renewbibmacro{in:}{}
\renewbibmacro*{volume+number+eid}{%
  \printfield{volume}
  \setunit*{\addnbthinspace}
  \printfield{number}
  \setunit{\addcomma\space}
  \printfield{eid}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\AtEveryBibitem{
  \clearfield{issn}
  \clearfield{month}
  \clearfield{urlyear}
  \clearlist{language}
  \clearfield{note}
  \ifentrytype{online}{}{
    \clearfield{url}
  }
}
\addbibresource{ForeignAidBib.bib}
\begin{document}
\maketitle
\section*{Introduction}
The People's Republic of China (China) established the Belt and Road Initiative (BRI) in 2013, a project aimed at increasing both China's global influence and domestic economic growth. Since then, China has become the ``lender of first resort'' for developing countries \parencite[1]{dreher2022}.

There has been an abundance of scholarly work attempting to understand the potential effects of China's recent prominence in the international financing and aid arena. Most, if not all, of this literature focuses on how the behavior of other aid donors has been affected (e.g., \cite{humphrey2019}; \cite{kilama2016a}) or how the effectiveness of their aid has changed (e.g., \cite{blair2022}; \cite{gehring2022}). Essentially, this work has explored how the supply side of foreign aid has been influenced by China's aid involvement.

However, there has been little research on how Chinese aid affects the demand side of foreign aid (i.e., recipient behavior), despite the supply side literature implying that the presence of Chinese aid is indeed consequential for recipients. In particular, it appears that the democratizing effects of aid from the Organization for Economic Cooperation and Development (OECD) has been reduced in light of Chinese aid (\cite{li2017a}) and that the World Bank provides fewer conditions for it's loans to potentially compete with China (\cite{hernandez2017}).

Given this seemingly enhanced foreign aid competition, I argue that recipient behavior is no longer a trivial question, as the conventional wisdom assumes. It should be plausible that recipients engage in some strategy to maximize the financial package they receive from donors, possibly diversifying their aid diet to include Chinese funds and gain leverage on Western donors.

To explore the possible strategic behavior of international finance recipients in this situation, I build a formal bargaining model below. The model attempts to capture the flexibility of Chinese lending practices and the results of Chinese competition with more traditional, OECD donors.

Empirically, I focus on evaluating the model's implications in Africa using available data on Chinese and Western loans. Data availability for Chinese aid limits the period of interest from 2000-2017.

\section*{Theory}
\subsection*{Model}
In an attempt to model the effects of having more than one donor on recipient behavior, especially when donors display heterogeneity in their loaning practices, I formalize a bargaining interaction between a recipient country and two types of donors. Figure 1 illustrates the extensive-form of the interaction and can be found in the appendix.

@@

\subsection*{Implications}
The first donor's offer is largest, though, when the severity of recipient need is low (high $\delta$) and the second donor is anticipated to implement clear conditions (high $\gamma$). This implies the comparative statics that as severity of need increases (i.e., $\delta$ gets smaller) or clarity of conditionality decreases (i.e., $\gamma$ gets smaller), loan value decreases for the recipient. This analysis produces the following hypotheses, conditional on the presence of a diverse set of donors:
\begin{align*}
    &\textbf{Recipient Need is Inversely Related to Loan Value}\\
    H_{1_a}&:\;\text{An increase in severity of recipient need will decrease the amount of money offered via loans}\\
    H_{1_b}&:\;\text{A decrease in severity of recipient need will increase the amount of money offered via loans}\\
    &\textbf{Condition Clarity is Directly Related to Loan Value}\\
    H_{2_a}&:\;\text{An increase in the clarity of loan conditionality will increase the amount of money offered via loans}\\
    H_{2_b}&:\;\text{A decrease in the clarity of loan conditionality will decrease the amount of money offered via loans}\\
\end{align*}
In other words, I expect to find the following relationship:

\subsection*{Table 1}
\begin{tabular}{c|c c }
    & Severity of Recipient Need & Clarity of Loan Conditionality\\
    \hline
    High & Low Loan Value & High Loan Value\\
    Low & High Loan Value & Low Loan Value\\
\end{tabular}

\section*{Empirics}
\subsection*{Operationalization}
To evaluate the theoretical expectations I focus on the international loan environment in Africa since the year 2000. Loan value, or the total sum of money being loaned to an African country, is the dependent variable. The main explanatory variables of interest are recipient need and the clarity of loan conditionality. Recipient need is captured by gross domestic product (GDP) per capita while the clarity of loan conditionality will be an indicator for how ``offical'' aid is. Thus, the empirical model is:

$\text{Loan Value}= \beta_0 + \beta_1 \text{GDP per capita}+ \beta_2 \text{Loan Conditionality}+ \gamma_i x_i + \varepsilon$,

where $x_i$ is a vector of controls.

\subsection*{Data \& Variables}
For data on international loans, I use data from William \& Mary's ``Aiddata'' lab for Chinese loans (\cite{custer2021}) and data from the OECD on Western loans (\cite{oecd2022}). GDP per capita stems from the World Bank (\cite{bank2022}) while I am currently exploring measures for loan conditionality. The most overly-simplistic (and not accurate) idea is to label all Chinese loans as ``low'' or ``vague'' in their conditionality while rating all Western loans as ``high'' or ``clear'' in their loan conditions. There is data on World Bank conditionality, though, (see \cite{hernandez2017}, for example) but I have yet to analyze the Chinese data for conditionality measures.

\subsection*{Estimation}
I plan to estimate the regression equation above using panel data and OLS with country-fixed effects and clustered standard errors by country.
